---
title: "Toronto 2023 Monthly Death Registry"
author: Yunzhao Li
thanks: "Code and data are available at: [https://github.com/yunzhaol/Toronto_Death_Registry.git](https://github.com/yunzhaol/Toronto_Death_Registry.git)."
date: today
date-format: long
abstract: "In Toronto, there are many people who die each year. The data of death registry supports the City's operational requirements and business functions. In this paper we will make a table of the number of death in Toronto for each month in 2023. By analyzing this, we can find months with a high number of deaths and make conjectures with factors like weather, policies or incidents of that particular month."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---


```{r}
#| echo: false
#| warning: false
#| message: false

#### Workspace set-up ####
library(tidyverse)
library(janitor)
library(opendatatoronto)
library(lubridate)
library(knitr)
library(lintr)
library(renv)
library(styler)
```


# Introduction

Death is a topic that people are not often willing to talk about, but it is closely related to the City's operational requirements, business functions as well as social resource allocation, so it is necessary to understand the data of death registration and do some basic statistical analysis of it. 

The data source in this study was obtained by opendatatoronto, from which the data of the Toronto area in 2023 were selected for research and analysis. We expect to produce a table reflecting the number of deaths registered in the Toronto area in each month of 2023, and correlate the data with related factors.

These speculation may help in the formulation of future policies, such as whether the death toll is higher in the colder months, then the government can increase investment in the construction of shelters, if it is because of the epidemic or epidemic in certain months, then it can be prevented according to the relevant month and transmission trend, formulate a mask-related policy or appropriately restrict travel, or because of some large-scale accident that caused a large number of new deaths. Which months have a low number of deaths can also be related to the input of some public resources, the increase in per capita income, and the improvement of social welfare, which in turn can be used to judge whether these policies and improvements are significant, whether people's quality of life and income are stable, and whether social security is good. The government can regulate these factors to reduce the number of deaths caused by non-aging, which is the long-term significance of this study. 

This study will end with these numbers documenting the lives that have accompanied the city and telling their stories implicitly to some extent.


# Data

This dataset includes information relating to the registration of deaths documented by Registry Services staff located in four of the civic centres (Scarborough, North York, Toronto and Etobicoke). Registration of death data is entered into the Registry Services Tracking System (RSTS), from which aggregate statistical information is generated to create the dataset. The dataset resides in an Oracle database in the City's environment.

The dataset is created in support of the Vital Statistics Act, which is a Provincial legislation. The dataset also supports the City's operational requirements and business functions.
[ https://open.toronto.ca/dataset/death-registry-statistics/]("https://open.toronto.ca/dataset/death-registry-statistics/")

The original data has five columns, and the column name and introduction are as followed:

row: identifier for Open Data database,

CIVIC_CENTRE:	Civic centre code,

DEATH_LICENSES:	Number of deaths registered in the month,

PLACE_OF_DEATH:	Place where death occurred,

TIME_PERIOD:	Month death registered.
Their classes are separately: numeric, character, numeric, character and numeric.

Depending on our expectation, only two columns named time_period and death_license after filtering by region and time period are needed. This column of data, called the time period, is presented in the form of yyyy-mm characters. After that, we need to modify it to date and count by month. The column, titled Death Certificate, is presented in numeric form, with a maximum value of 2007 and a minimum value of 1, reflecting a single record of local deaths over each time period. And for the column called place_of_death, can be used to filter out death records in the Toronto area. 

This is the basic description of the data of interests. The appendix shows the process of simulating, reading, cleaning, and creating tables of data in details.


```{r}
#| echo: false
#| warning: false
#| message: false
#### Acquire ####
# Download data 
raw_death_registry <-
  read_csv(
    here::here("inputs/data/Death Registry Statistics Data.csv"),
    show_col_types = FALSE
  )

# Save data
write_csv(
  x = raw_death_registry,
  file = "raw_death_registry.csv"
)
```



```{r}
#| echo: false
#| warning: false
#### Clean ####
# filter the death registry in Toronto and in 2023
cleaned_death_registry <-
  clean_names(raw_death_registry) |>
  mutate(time_period = ym(time_period)) |>
  filter(
    place_of_death == "Toronto",
    time_period %in% c(sprintf("2023-0%s-01", c(1:9)), "2023-10-01", "2023-11-01", "2023-12-01")
  ) |>
  select(time_period, death_licenses)
```



```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#### Test ####
# check the class and value, death_licenses must be greater than 1 and less than 2007.
cleaned_death_registry$time_period |>
  unique() == c(sprintf("2023-0%s-01", c(1:9)), "2023-10-01", "2023-11-01", "2023-12-01")
cleaned_death_registry$time_period |> class() == "Date"
cleaned_death_registry$death_licenses |> class() == "numeric"
cleaned_death_registry$death_licenses |> min() >= 1
cleaned_death_registry$death_licenses |> max() <= 2007
```



```{r}
#| echo: false
#| warning: false
# Save data
write_csv(
  x = cleaned_death_registry,
  file = "cleaned_death_registry.csv"
)
```


```{r}
#| echo: false
#| warning: false
#### Explore ####
# read in clean data
cleaned_death_registry <-
  read_csv(
    "cleaned_death_registry.csv",
    show_col_types = FALSE
  )
```

\newpage

# Results

To create the table we expect follows this procedure: simulate with a dataset, read the raw data, clean the data, and finally the data was accurately presented and consistent with the sketch we made. The number of deaths in each month is as followed:

January: 1083, February: 485, March: 1012, April: 1677, May: 2008, June: 1928, July: 417, August: 912, September: 1268, October: 1182, November: 1458, December: 611. 

We found that the highest number of deaths was in May and June, with a high of around 2,000, and the lowest months were in February, July and December, with around 500. The death toll in the remaining months was concentrated around 1,000. 

Combined with the ideas introduced at the beginning of the article, this data can be combined with other relevant data to analyze, and perhaps unexpected discoveries can be obtained. This is the direction of future research that can be further explored, and this data can be used as a basis. You can also drill down into other years' data, or trends in total deaths over several years, to create a line chart.

The result of our code are shown by @tbl-deathregistrytable.
```{r}
#| echo: false
#| warning: false
#| label: tbl-deathregistrytable
#| tbl-cap: "Number of death in Toronto for each month in 2023"

# create a table of death registry for each month
cleaned_death_registry |>
  mutate(time_period_month = month(
    time_period,
    label = TRUE,
    abbr = FALSE
  )) |>
  arrange(time_period_month) |>
  drop_na(death_licenses) |>
  summarise(
    death_licenses = sum(death_licenses),
    .by = time_period_month
  ) |>
  kable(
    col.names = c("Month", "Number of death")
  )

```

\newpage

# Conclusion
Overall, this study successfully plotted the number of deaths registered in Toronto by month in 2023, and achieved the expected results. The highest number of deaths was concentrated in May and June, with a high of around 2,000, and the lowest months were in February, July and December, with a low of around 500. The death toll in the remaining months was concentrated around 1,000. It is believed that data can be used to assist in government management, policy formulation, and improvement of people's livelihood.

\newpage

\appendix

# Appendix {-}

# Additional data details

## Simulation

We simulated the dataset using the tipple function and determined the feasibility of the plan。

The top 6 rows are shown by @tbl-simulatedata

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-simulatedata
#| tbl-cap: "Dataset Simulation"
#### Simulate ####
# simulate the dataset
set.seed(23)

simulated_death_data <-
  tibble(
    time_period = rep(c(sprintf("2023-0%s", c(1:9)), sprintf("2023-1%s", c(0:2))), times = 3),
    place_of_death = c(rep("Toronto", times = 36)),
    death_number =
      rpois(
        n = 36,
        lambda = 50
      ) # Draw 36 times from the Poisson distribution
  )
# create a table to see the simulated data of top 6 rows 
head(simulated_death_data) |>
  kable(
    col.names = c("time period", "place_of_death", "death_number"),
    booktabs = TRUE
  )
```


## Read raw data

We downloaded a csv document which documents the data of this topic: About Death Registry Statistics from opendatatoronto as raw data.

The top 6 rows are shown by @tbl-rawdata

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-rawdata
#| tbl-cap: "Raw Death Registry"
#### Acquire ####
# create a table to see the raw data of top 6 rows 
head(raw_death_registry) |>
  kable(
    col.names = c("_id", "CIVIC_CENTRE", "DEATH_LICENSES", "PLACE_OF_DEATH", "TIME_PERIOD"),
    booktabs = TRUE
  )
```



## Clean data

The purpose of this step is to filter out the two columns of data we need for the Toronto area and the time period in 2023, in preparation for the table we about to create. First, the clean_names function is used to change the column name to lowercase and use _ to split the word. Then the date character type is converted to the date type. After that, we filtered the data of the death register in Toronto and belonging to 2023 by listing all possible month data. Finally, select the two column we need, which is time_period and number of death_license.

The top 6 rows are shown by @tbl-cleanedata
```{r}
#| echo: false
#| warning: false
#| label: tbl-cleanedata
#| tbl-cap: "Cleaned Death Registry"
#### Clean ####
# create a table to see the cleaned data of top 6 rows 
head(cleaned_death_registry) |>
  kable(
    col.names = c("time_period", "death_licenses"),
    booktabs = TRUE
  )
```

## Tests

Three tests were wrote down to check whether the cleaned death registry data suits our expectation. Separately test time_period's value, time_period's class and death_license's class.
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#### Test ####
# check the class and value, death_licenses must be greater than 1 and less than 2007.
cleaned_death_registry$time_period |>
  unique() == c(sprintf("2023-0%s-01", c(1:9)), "2023-10-01", "2023-11-01", "2023-12-01")
cleaned_death_registry$time_period |> class() == "Date"
cleaned_death_registry$death_licenses |> class() == "numeric"
cleaned_death_registry$death_licenses |> min() >= 1
cleaned_death_registry$death_licenses |> max() <= 2007
```




\newpage


# References
@citeR
@tellingstories
@citeopendatatoronto
@citetidyverse
@citeR-knitr
@citelubridate
@citejanitor
@citerenv
@citelintr
@citestyler
@Inessa
