---
title: "Death Registry "
author: Yunzhao Li
thanks: "Code and data are available at: LINK."
date: 21 January 2024
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
bibliography: references.bib
---

```{r}
#| echo: false
#| warning: false
install.packages("tidyverse")
install.packages("janitor")
install.packages("opendatatoronto")
install.packages("lubridate")
install.packages("knitr")
```

```{r}
#| include: true
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(opendatatoronto)
library(lubridate)
library(knitr)

#### Acquire ####
raw_death_registry <-
  read_csv(
    file =
      "Death Registry Statistics Data.csv",
    show_col_types = FALSE
)

write_csv(
  x = raw_death_registry,
  file = "raw_death_registry.csv"
)

head(raw_death_registry)
```

```{r}
#### Clean ####
cleaned_death_registry <-
  clean_names(raw_death_registry) |>
  mutate(time_period = ym(time_period)) |> 
  filter(
    place_of_death == "Toronto",
    time_period %in% c("2023-01-01","2023-02-01","2023-03-01","2023-04-01","2023-05-01","2023-06-01","2023-07-01","2023-08-01","2023-09-01","2023-10-01","2023-11-01","2023-12-01")
    ) |>
  select(time_period, death_licenses)

head(cleaned_death_registry)
```

```{r}
write_csv(
  x = cleaned_death_registry,
  file = "cleaned_death_registry.csv"
)
```


```{r}
#### Explore ####
cleaned_death_registry <-
  read_csv(
    "cleaned_death_registry.csv",
    show_col_types = FALSE
  )
```


```{r}
cleaned_death_registry |>
  mutate(time_period_month = month(
    time_period,
    label = TRUE,
    abbr = FALSE
  )) |>
  arrange(time_period_month) |> 
  drop_na(death_licenses) |>
  summarise(death_licenses = sum(death_licenses),
            .by = time_period_month) |>
  kable(
    col.names = c("Month", "Total number of death in Toronto this month")
  )

```





# Introduction

You can and should cross-reference sections and sub-sections. 

The remainder of this paper is structured as follows. @sec-data....



# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 



# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


